### 2021-02-14 _post/2021-02-12-thankyoublogdownhellodistill/thankyoublogdownhellodistill.Rmd

---
title: "Thank you blogdown, hello distill"
description: |
  A short description of the post.
author:
  - name: Roland Schmidt
    url: https://werk.statt.codes
date: 02-12-2021

draft: T

output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

So, I moved my collection of posts ('blog' sounds a bit to serious) from blogdown to distill. 
In principle, setting up the new blog (damn...) was quite straighforward, particularly thanks to the
comprehensive documentation on the distill site and posts by Thomas, XXXX and XXXX.

Nevertheless, a few quick remarks which might be helpful to those who are also considerng to move:

Search function:
  Not that the search bar will only show up once you've deployed your site (e.g. via githut to netlify). 
  I initially was trying to find an error until I mustered the courage to post a pertaining question on github, and
  Jeff was kind enough to quickly respond. The mentioned 'sevr' package gives you a handy tool to test-drive your
  site before deploying.

Where to put static files:
_blog_data

theming:
css

performance:
tick slower than blogdown / based on hugo