### 2021-02-17 _post/2021-02-17-ggiraphtext/ggiraphtext.Rmd
add ggiraph test

---
title: "ggiraph text test"
description: |
  A short description of the post.
author:
  - name: Roland Schmidt
    url: https://werk.statt.codes
date: 02-17-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r}
library(ggplot2)
library(ggiraph)
library(svglite)
library(ggtext)
data <- mtcars
data$carname <- row.names(data)

```

# Normal, theme_minmal

```{r}
gg_point = ggplot(data = data) +
  labs(title="long title long title long title")+
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = carname, data_id = carname)) +
  theme_minimal()

girafe(ggobj = gg_point)
```

# Normal, theme_minmal, plot_title_elmenent_markdown

```{r}
gg_point = ggplot(data = data) +
  labs(title="long title long title long title")+
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = carname, data_id = carname)) +
  theme_minimal()+
  theme(plot.title=element_markdown())

girafe(ggobj = gg_point)
```

# Normal, theme_minmal, plot_title_elmenent_markdown, device svglite

```{r, dev="svglite"}
gg_point = ggplot(data = data) +
  labs(title="long title long title long title")+
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = carname, data_id = carname)) +
  theme_minimal()+
  theme(plot.title=element_markdown())

girafe(ggobj = gg_point)
```